NGO Repression as a Predictor of Worsening Human Rights Abuses

### 2021-11-12 manuscript/manuscript.Rmd
Fix issue 13 from R2: trichotomize the PTS

In general, when holding all other covariates constant, the passage of an additional anti-NGO legal barrier is generally associated with a higher probability of more severe level of political terror in the following year. The probability of seeing different levels of political terror varies across different possible counts of NGO laws (see Figure \@ref(fig:plot-mfx-e1a)). In the absence of formal legal barriers, Level 2 of the PTS—an environment where political imprisonment, torture, and beatings are rare—is the most likely category of political terror across all models, appearing a predicted 60% of the time. Level 1, where a country is under secure rule of law, is the second most likely outcome at slightly under 40%. The other categories of the PTS are extremely rare when there are no anti-NGO laws.

```{r plot-mfx-e1a, fig.width=6.5, fig.height=5, fig.asp=NULL, fig.cap="Marginal effects of increasing anti-NGO legal barriers on the probability of specific levels of political terror"}
mfx_e1a_pts %>% 
  mutate(x_rounded = floor(effect1__)) %>% 
  group_by(x_rounded, effect2__, plot_var_nice) %>% 
  slice(1) %>% 
  ungroup() %>% 
  ggplot(aes(x = x_rounded, y = estimate__, color = effect2__)) +
  geom_line(aes(group = effect2__), size = 0.5, color = "grey90") +
  geom_pointrange(aes(ymin = lower__, ymax = upper__),
                  position = position_dodge(width = 0.2),
                  fatten = 1, size = 1) +
  scale_x_continuous(breaks = 0:9) +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  scale_color_viridis_d(option = "rocket", end = 0.85, direction = -1) +
  guides(color = guide_legend(override.aes = list(size = 0.5, linetype = 0))) +
  labs(x = "Count of legal barriers\n",
       y = "Predicted probability of category",
       color = "Predicted Political Terror Scale", fill = "Predicted Political Terror Scale",
       title = "Political Terror Scale") + 
  facet_wrap(vars(plot_var_nice), scales = "free_x") +
  theme_ngo() +
  theme(panel.grid.major.x = element_blank())
```

As new laws are added, though, the distribution of probabilities for each of the PTS levels shifts. Adding new barriers to advocacy, entry, and funding decreases the likelihood of seeing Level 1 and increases the probability of a country turning to either Level 2 or Level 3 (with extensive political imprisonment and trial-free detention). The marginal effects of all forms of NGO laws results in this turn from Level 1 to Levels 2 or 3, but this switching effect is strongest following additional barriers to advocacy. Anti-NGO laws do not predict more dramatic shifts in the PTS. Across all counts of legal barriers, the overall probability of seeing Levels 4 and 5 is minuscule, and the passave of additional laws does little to shift that probability. This is not unexpected however—the PTS has a natural ceiling at Level 5. Specifically, in addition to measuring the scope and intensity of government violence, the PTS scale also looks at the proportion of population targeted for abuse. Countries that receive a 4 and 5 are rare, as 4 denotes that murders, disappearances, and torture are a common part of life, while 5 denotes that physical abuses are routinely perpetrated against citizens not involved in politics, and includes cases where governments engage in "scorched earth" counterinsurgency policies or summary executions based on class or ethnic affiliation [@GibneyCornettWood:2019]. The majority of movement across PTS categories should thus occur between Levels 1–3.

@@

```{r plot-mfx-e1b-e1d, fig.width=7.5, fig.height=7.5, fig.asp=NULL, fig.cap="Marginal effects of increasing anti-NGO legal barriers on V-Dem rights indexes and latent human rights"}
out_mfx_e1b <- mfx_e1b_clphy %>% 
  mutate(x_rounded = floor(effect1__)) %>% 
  group_by(x_rounded, plot_var_nice) %>% 
  slice(1) %>% 
  ungroup() %>% 
  mutate(plot_var_nice = fct_relabel(plot_var_nice, ~str_to_sentence(str_remove(., "Barriers to ")))) %>% 
  ggplot(aes(x = x_rounded, y = estimate__, color = plot_var_nice)) +
  geom_line(size = 0.5, color = "grey90") +
  geom_pointrange(aes(ymin = lower__, ymax = upper__),
                  fatten = 1, size = 1) +
  scale_x_continuous(breaks = 0:9) +
  scale_color_viridis_d(option = "plasma", end = 0.8) +
  coord_cartesian(ylim = c(0.60, 0.71)) +
  guides(color = "none") +
  labs(x = NULL,
       y = "Predicted physical violence index",
       title = "Physical violence index",
       caption = "(Higher values = less violence)") +
  facet_grid(cols = vars(plot_var_nice), scales = "free_x", space = "free_x") +
  theme_ngo() +
  theme(panel.grid.major.x = element_blank())

out_mfx_e1c <- mfx_e1c_clpriv %>% 
  mutate(x_rounded = floor(effect1__)) %>% 
  group_by(x_rounded, plot_var_nice) %>% 
  slice(1) %>% 
  ungroup() %>% 
  mutate(plot_var_nice = fct_relabel(plot_var_nice, ~str_to_sentence(str_remove(., "Barriers to ")))) %>% 
  ggplot(aes(x = x_rounded, y = estimate__, color = plot_var_nice)) +
  geom_line(size = 0.5, color = "grey90") +
  geom_pointrange(aes(ymin = lower__, ymax = upper__),
                  fatten = 1, size = 1) +
  scale_x_continuous(breaks = 0:9) +
  scale_color_viridis_d(option = "plasma", end = 0.8) +
  coord_cartesian(ylim = c(0.60, 0.71)) +
  guides(color = "none") +
  labs(x = NULL,
       y = "Predicted civil liberties index",
       title = "Private civil liberties index",
       caption = "(Higher values = better protection)") +
  facet_grid(cols = vars(plot_var_nice), scales = "free_x", space = "free_x") +
  theme_ngo() +
  theme(panel.grid.major.x = element_blank())

out_mfx_e1d <- mfx_e1d_lhr %>% 
  mutate(x_rounded = floor(effect1__)) %>% 
  group_by(x_rounded, plot_var_nice) %>% 
  slice(1) %>% 
  ungroup() %>% 
  mutate(plot_var_nice = fct_relabel(plot_var_nice, ~str_to_sentence(str_remove(., "Barriers to ")))) %>% 
  ggplot(aes(x = x_rounded, y = estimate__, color = plot_var_nice)) +
  geom_line(size = 0.5, color = "grey90") +
  geom_pointrange(aes(ymin = lower__, ymax = upper__),
                  fatten = 1, size = 1) +
  scale_x_continuous(breaks = 0:9) +
  scale_color_viridis_d(option = "plasma", end = 0.8) +
  guides(color = "none") +
  labs(x = "Count of legal barriers",
       y = "Predicted latent human rights value",
       title = "Latent human rights",
       caption = "(Higher values = better protection)") +
  facet_grid(cols = vars(plot_var_nice), scales = "free_x", space = "free_x") +
  theme_ngo() +
  theme(panel.grid.major.x = element_blank())

out_mfx_e1b / out_mfx_e1c / out_mfx_e1d
```

@@

The effect of de facto civil society repression on predicted values of the PTS mirrors what we found previously with our first empirical expectation. As seen in Figure \@ref(fig:plot-mfx-e2a), at the lowest levels of repression, Levels 1 and 2 are the most common predicted outcomes and are each equiprobable at roughly 50%. As de facto CSO repression increases, though, the probability of seeing Level 1 decreases and is replaced with either Level 2 or Level 3. More severe levels of political terror remain exceptionally improbable at any level of preceding civil society repression.

```{r plot-mfx-e2a, fig.width=6.5, fig.height=5, fig.asp=NULL, fig.cap="Marginal effects of changing levels of civil society repression on the probability of specific levels of political terror", out.width="75%"}
out_mfx_e2a <- mfx_e2a_pts %>% 
  ggplot(aes(x = effect1__, y = estimate__, color = effect2__, fill = effect2__)) +
  geom_ribbon(aes(ymin = lower__, ymax = upper__), alpha = 0.4, color = NA) +
  geom_line(size = 1) +
  scale_x_reverse() +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  scale_color_viridis_d(option = "rocket", end = 0.75, direction = -1) +
  scale_fill_viridis_d(option = "rocket", end = 0.9, direction = -1) +
  labs(x = "Civil society repression",
       y = "Predicted probability of category",
       color = "Political Terror Scale", fill = "Political Terror Scale",
       title = "Political Terror Scale",
       caption = "(x-axis reversed; greater repression on the right)") +
  theme_ngo() +
  theme(legend.box.margin = margin(t = 0.5, unit = "lines"))
out_mfx_e2a
```

>>>>

In general, when holding all other covariates constant, the passage of an additional anti-NGO legal barrier is generally associated with a higher probability of more severe level of political terror in the following year. The probability of seeing different levels of political terror varies across different possible counts of NGO laws (see Figure \@ref(fig:plot-mfx-e1a)). In the absence of formal legal barriers, Level 2 of the PTS—an environment where political imprisonment, torture, and beatings are rare—is the most likely category of political terror across all models, appearing a predicted 60% of the time. Level 1, where a country is under secure rule of law, is the second most likely outcome at slightly under 40%. The other categories of the PTS are extremely rare when there are no anti-NGO laws—as such, we collapse them into a single category.

```{r plot-mfx-e1a, fig.width=6.5, fig.height=5, fig.asp=NULL, fig.cap="Marginal effects of increasing anti-NGO legal barriers on the probability of specific levels of political terror"}
mfx_e1a_pts %>% 
  mutate(x_rounded = floor(effect1__)) %>% 
  # Collapse 3-5 to just one level, since they're all tiny
  mutate(effect2__ = fct_collapse(effect2__, "3+" = c("3", "4", "5"))) %>% 
  group_by(x_rounded, effect2__, plot_var_nice) %>% 
  slice(1) %>% 
  # Add up these values for the collapsed 3-5 results
  summarize(across(c(estimate__, lower__, upper__), sum)) %>% 
  ungroup() %>% 
  ggplot(aes(x = x_rounded, y = estimate__, color = effect2__)) +
  geom_line(aes(group = effect2__), size = 0.5, color = "grey90") +
  geom_pointrange(aes(ymin = lower__, ymax = upper__),
                  # position = position_dodge(width = 0.2),
                  fatten = 1, size = 1) +
  scale_x_continuous(breaks = 0:9) +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  scale_color_viridis_d(option = "rocket", begin = 0.1, end = 0.85, direction = -1) +
  guides(color = guide_legend(override.aes = list(size = 0.5, linetype = 0))) +
  labs(x = "Count of legal barriers\n",
       y = "Predicted probability of category",
       color = "Predicted Political Terror Scale", fill = "Predicted Political Terror Scale",
       title = "Political Terror Scale") + 
  facet_wrap(vars(plot_var_nice), scales = "free_x") +
  theme_ngo() +
  theme(panel.grid.major.x = element_blank())
```

As new laws are added, though, the distribution of probabilities for each of the PTS levels shifts. Adding new barriers to advocacy, entry, and funding decreases the likelihood of seeing Level 1 and increases the probability of a country turning to either Level 2 or Level 3 (with extensive political imprisonment and trial-free detention). The marginal effects of all forms of NGO laws results in this turn from Level 1 to Levels 2 or 3+, but this switching effect is strongest following additional barriers to advocacy. Anti-NGO laws do not predict more dramatic shifts in the PTS. Across all counts of legal barriers, the overall probability of seeing Levels 4 and 5 is minuscule, and the passage of additional laws does little to shift that probability. This is not unexpected however—the PTS has a natural ceiling at Level 5. Specifically, in addition to measuring the scope and intensity of government violence, the PTS scale also looks at the proportion of population targeted for abuse. Countries that receive a 4 and 5 are rare, as 4 denotes that murders, disappearances, and torture are a common part of life, while 5 denotes that physical abuses are routinely perpetrated against citizens not involved in politics, and includes cases where governments engage in "scorched earth" counterinsurgency policies or summary executions based on class or ethnic affiliation [@GibneyCornettWood:2019]. The majority of movement across PTS categories should thus occur between Levels 1–3.

@@

```{r plot-mfx-e1b-e1d, fig.width=7.5, fig.height=4, fig.asp=NULL, fig.cap="Marginal effects of increasing anti-NGO legal barriers on V-Dem rights indexes and latent human rights"}
out_mfx_e1d <- mfx_e1d_lhr %>% 
  mutate(x_rounded = floor(effect1__)) %>% 
  group_by(x_rounded, plot_var_nice) %>% 
  slice(1) %>% 
  ungroup() %>% 
  mutate(plot_var_nice = fct_relabel(
    plot_var_nice, ~str_to_sentence(str_remove(., "Barriers to ")))) %>% 
  ggplot(aes(x = x_rounded, y = estimate__, color = plot_var_nice)) +
  geom_line(size = 0.5, color = "grey90") +
  geom_pointrange(aes(ymin = lower__, ymax = upper__),
                  fatten = 1, size = 1) +
  scale_x_continuous(breaks = 0:9) +
  scale_color_viridis_d(option = "plasma", end = 0.8) +
  guides(color = "none") +
  labs(x = "Count of legal barriers",
       y = "Predicted latent human rights value",
       title = "Latent human rights",
       caption = "(Higher values = better protection)") +
  facet_grid(cols = vars(plot_var_nice), scales = "free_x", space = "free_x") +
  theme_ngo() +
  theme(panel.grid.major.x = element_blank())

out_mfx_e1d
```

@@

The effect of de facto civil society repression on predicted values of the PTS mirrors what we found previously with our first empirical expectation. As seen in Figure \@ref(fig:plot-mfx-e2a), at the lowest levels of repression, Levels 1 and 2 are the most common predicted outcomes and are each equiprobable at roughly 50%. As de facto CSO repression increases, though, the probability of seeing Level 1 decreases and is replaced with either Level 2 or Level 3. More severe levels of political terror remain exceptionally improbable at any level of preceding civil society repression—we again collapse these into a single category.

```{r plot-mfx-e2a, fig.width=6.5, fig.height=5, fig.asp=NULL, fig.cap="Marginal effects of changing levels of civil society repression on the probability of specific levels of political terror", out.width="75%"}
out_mfx_e2a <- mfx_e2a_pts %>% 
  mutate(effect2__ = fct_collapse(effect2__, "3+" = c("3", "4", "5"))) %>% 
  group_by(effect1__, effect2__, plot_var_nice) %>% 
  summarize(across(c(estimate__, lower__, upper__), sum)) %>% 
  ungroup() %>% 
  ggplot(aes(x = effect1__, y = estimate__, color = effect2__, fill = effect2__)) +
  geom_ribbon(aes(ymin = lower__, ymax = upper__), alpha = 0.4, color = NA) +
  geom_line(size = 1) +
  scale_x_reverse() +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  scale_color_viridis_d(option = "rocket", begin = 0.1, end = 0.75, direction = -1) +
  scale_fill_viridis_d(option = "rocket", begin = 0.1, end = 0.9, direction = -1) +
  labs(x = "Civil society repression",
       y = "Predicted probability of category",
       color = "Political Terror Scale", fill = "Political Terror Scale",
       title = "Political Terror Scale",
       caption = "(x-axis reversed; greater repression on the right)") +
  theme_ngo() +
  theme(legend.box.margin = margin(t = 0.5, unit = "lines"))
out_mfx_e2a
```
